---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "70%"
)
```

# crimtools

<!-- badges: start -->
<!-- badges: end -->

`crimtools` is a collection of packages intended to make it easier for researchers
to perform common data analyses on spatial data. This package interfaces with
spatial data in `sf` format and provides functions to help plotting 
static maps via `ggplot2` or interactive maps via `maptools`. This is also 
intended to serve as a centralized point for basic spatial analyses as many of
the previous packages are no longer being updated (such as `GISTools`) or are
being depricated in favor of `sf`, `stars` and `terra` (see [here](https://r-spatial.org//r/2022/04/12/evolution.html) ). The core
functionality of `crimtools` is based around either raster or areal analyses. 
While this package is still in active development, the intended main functions
will be:

1. Kernel density mapping
2. Spatial grids and grid counts
3. LISA statistics (Moran's I, Getis Ord Gi*)
  
In the future, other functionalities provided in commercial tools (i.e. ArcGIS)
will be ported to `crimtools`.

## Installation

You can install the development version of crimtools from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("gmcirco/crimtools")
```
## Tools Available

The `crimtools` package is intended to be used with simple features from the `sf`
package (see [this link](https://r-spatial.github.io/sf/articles/sf1.html) for a 
more in-depth description of simple features in R). Currently, the following 
analyses are supported:

1. [Kernel density estimation](#kde)
2. [DBCAN point clustering](#dbs)

## Kernel Density Estimation{#kde}

The `kernel_density` function requires, at a minimum, a point feature `x` and
an enclosing polygon feature `region`. Both of these should be `sf` objects 
and should (ideally) be projected into feet or meters. By default, `kernel_density`
will automatically estimate a bandwidth by using the `bw.ppl` function as part
of `spatstat.core`. However, users can (and should) provide a bandwidth size that
reflects domain knowledge about the relevant feature. Users can also provide 
a custom function to estimate the bandwidth, or use several of those available in
`spatstat`.

```{r kde1}
# Load necessary libraries
library(crimtools)
library(sf)

# Load example data
data("newhaven")
data("nh_hom")

# Default output, gaussian kernel
kde_out <- kernel_density(x = nh_hom, region = newhaven)
```

The kernel type, bandwidth, and other features can be manually adjusted as well.

```{r}
# Specifying a quartic kernel with 2000 foot bandwidth
kde_out2 <- kernel_density(x = nh_hom, 
                           region = newhaven, 
                           kernel = "quartic", 
                           bdw = 2000)
```

By default `kernel_density` returns a dataframe with columns for the density value and
the x and y coordinates of cell centroid. 

```{r}
head(kde_out)
```

The exported dataframe can easily be plotted using `ggplot2` functionality, or 
used as a variable in an analysis. This allows for easy customization using the
large number of options available in `ggplot2`

```{r kde2}
library(ggplot2)

# plot as a ggplot2 object using geom_tile
# and setting coord_equal
ggplot(kde_out) +
  geom_tile(aes(x = X, y = Y, fill = density)) +
  coord_equal() +
  scale_fill_viridis_c(option = "magma") +
  theme_void()

```

## DBSCAN Point Clustering {#kde}

Given a region and point pattern, this function computes clusters based on 
user-defined search parameters. Here, this is the `dist` measure and `minPts` 
measure. This means we will search for clusters that are within a 2000 foot radius
of a 'core' point, and contain a minimum of 5 points. 
